---
title: "Project_2"
author: "Rahul"
date: "2024-04-24"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
setwd("C:/Users/rahul/OneDrive/Desktop/Notes/SS-24/ICS/Project_2")

```
#Loadind and setups

```{r, echo = TRUE}

library(haven)
library(ggplot2)
library(dplyr)

dataset <- read_sav("Sandwich.sav")

# Renaming the covariates
names(dataset) <- c("Ants", "Bread", "Topping", "Butter")

dataset$Bread <- as.factor(dataset$Bread)
levels(dataset$Bread) <- c("wholegrain", "multigrain", "rye", "white bread")

dataset$Topping <- as.factor(dataset$Topping)
levels(dataset$Topping) <- c("ham pickles", "peanut butter", "yeast spread")

dataset$Butter <- as.factor(dataset$Butter)
levels(dataset$Butter) <- c ("with", "without")

names(dataset)
summary(dataset)
```

#Box Plots for each factor

```{r, echo = TRUE}
# Bread 
gg.Bread <- ggplot(dataset, aes(x = Bread, y = Ants)) +
  geom_boxplot(fill = "skyblue", color = "blue") +
  stat_summary(fun = mean, geom = "point", shape = 4, size = 3, 
               color = "black") +  # Add mean point
  labs(x = "", y = "Number of Ants"
       # , title = "Box Plot of Ants by Bread type"
       ) +
  theme_minimal()



# Toppings
gg.Topping <- ggplot(dataset, aes(x = Topping, y = Ants)) +
  geom_boxplot(fill = "lightgreen", color = "darkgreen") +
  stat_summary(fun = mean, geom = "point", shape = 4, size = 3, 
               color = "black") +  # Add mean point
  labs(x = "", y = "Number of Ants"
       # , title = "Box Plot of Ants by Topping type"
       ) +
  theme_minimal()



# Butter
gg.Butter <- ggplot(dataset, aes(x = Butter, y = Ants)) +
  geom_boxplot(fill = "lightpink", color = "red") +
  stat_summary(fun = mean, geom = "point", shape = 4, size = 3, 
               color = "black") +  # Add mean point
  labs(x = "", y = "Number of Ants"
       # , title = "Box Plot of Ants by with or without Butter"
       ) +
  theme_minimal()



gg.Bread; gg.Topping; gg.Butter 

```



# Test Assumptions
1 Equal variance among the groups.
2 The data is normally distributed
3 The data is independent.

#Statistical Summary for each factor
```{r, echo = TRUE}

#Bread : 4 levels
summary.Brot <- dataset %>% group_by(Bread) %>%
  summarise(
    count = n(),
    min = min(Ants, na.rm = TRUE),
    # Q1  = quantile(Ants,0.25),
    median = median(Ants, na.rm = TRUE),
    # Q3  = quantile(Ants,0.75),
    max  = max(Ants, na.rm = TRUE),
    mean = mean(Ants, na.rm = TRUE),
    sd = sd(Ants, na.rm = TRUE)
    # ,
    # IQR = quantile(Ants,0.75) - quantile(Ants,0.25)
  )
print(summary.Brot)

summary.Brot$sd

summary.Brot$sd[-2][3] - summary.Brot$sd[-2][2]


#Topping : 3 levels
summary.Topping <- dataset %>% group_by(Topping) %>%
  summarise(
    count = n(),
    min = min(Ants, na.rm = TRUE),
    # Q1  = quantile(Ants,0.25),
    median = median(Ants, na.rm = TRUE),
    # Q3  = quantile(Ants,0.75),
    max  = max(Ants, na.rm = TRUE),
    mean = mean(Ants, na.rm = TRUE),
    sd = sd(Ants, na.rm = TRUE)
    # ,
    # IQR = quantile(Ants,0.75) - quantile(Ants,0.25)
  )
print(summary.Topping)

summary.Topping$sd
max(summary.Topping$sd) - min(summary.Topping$sd)


#Butter : 2 levels
summary.Butter <- dataset %>% group_by(Butter) %>%
  summarise(
    count = n(),
    min = min(Ants, na.rm = TRUE),
    # Q1  = quantile(Ants,0.25),
    median = median(Ants, na.rm = TRUE),
    # Q3  = quantile(Ants,0.75),
    max  = max(Ants, na.rm = TRUE),
    mean = mean(Ants, na.rm = TRUE),
    sd = sd(Ants, na.rm = TRUE)
    # ,
    # IQR = quantile(Ants,0.75) - quantile(Ants,0.25)
  )
print(summary.Butter)

```

# QQ plot for normality test for each factor and their levels

```{r, echo = TRUE}
# Brot(Bread) QQ
qq_plot_Bread <- ggplot(dataset, aes(sample = Ants, color = Bread)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "QQ Plot : Ants by Bread type") +
  facet_wrap(~ Bread) +  # Create separate plots for each category of Bread
  theme_minimal() +  # Set theme to minimal
  theme(plot.title = element_text(size = 12), legend.position = "none") + # Decrease title font size
  xlab("Theoretical Quantiles") +  # Add x-axis label
  ylab("Sample Quantiles")




# Belag(Topping) QQ
qq_plot_Topping <- ggplot(dataset, aes(sample = Ants, color = Topping)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "QQ Plot : Ants by Topping type") +
  facet_wrap(~ Topping) +  # Create separate plots for each category of Belag
  theme_minimal() +  # Set theme to minimal
  theme(plot.title = element_text(size = 12), legend.position = "none")  + # Decrease title font size
  xlab("Theoretical Quantiles") +  # Add x-axis label
  ylab("Sample Quantiles")




# Butter QQ
qq_plot_Butter <- ggplot(dataset, aes(sample = Ants, color = Butter)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "QQ Plots : Ants by with or without Butter") +
  facet_wrap(~ Butter) +  # Create separate plots for each category of Butter
  theme_minimal() +  # Set theme to minimal
  theme(plot.title = element_text(size = 12), legend.position = "none") + # Decrease title font size
  xlab("Theoretical Quantiles") +  # Add x-axis label
  ylab("Sample Quantiles")


qq_plot_Bread; qq_plot_Topping; qq_plot_Butter
```


#ANOVA compares the mean in different group are equal or not
# ie. We will test for each factor with different levels 

```{r, echo = TRUE}

# Step 1: Load necessary libraries
library(stats)    # For statistical functions

#ANOVA - Bread
cat("Bread Group \n")
Bread_anova <- aov(Ants ~ Bread, data = dataset)
summary(Bread_anova)

summary_Bread_anova <- summary(Bread_anova)
summary_Bread_anova[[1]]$`Pr(>F)`[1]





#ANOVA - Topping
cat("\n Topping Group \n")
Topping_anova <- aov(Ants ~ Topping, data = dataset)
summary(Topping_anova)


# Summarize ANOVA results
summary_Topping_anova <- summary(Topping_anova)

# Extract p-value
summary_Topping_anova[[1]]$`Pr(>F)`[1]



```

## T-test for each factor with different levels 
# ie. pair-wise t test among different levels


```{r, echo = TRUE}

# Pairwise T-test for Topping ===================================

# To store p-values for each combination of levels of Topping
unadjusted.p.values <- c() 


## (1) "ham pickles"   &   (2) "peanut butter"
t.Topping.H01 <- t.test(dataset$Ants[dataset$Topping == "ham pickles"], 
       dataset$Ants[dataset$Topping == "peanut butter"], 
       alternative = "two.sided", paired = FALSE, var.equal = TRUE)

unadjusted.p.values[1] <- t.Topping.H01$p.value



## (1) "ham pickles"   &   (3) "yeast spread"
t.Topping.H02 <- t.test(dataset$Ants[dataset$Topping == "ham pickles"], 
       dataset$Ants[dataset$Topping == "yeast spread"], 
       alternative = "two.sided", paired = FALSE, var.equal = TRUE)

unadjusted.p.values[2] <- t.Topping.H02$p.value



## (2) "peanut butter"   &   (3) "yeast spread"
t.Topping.H03 <- t.test(dataset$Ants[dataset$Topping == "peanut butter"], 
       dataset$Ants[dataset$Topping == "yeast spread"], 
       alternative = "two.sided", paired = FALSE, var.equal = TRUE)

unadjusted.p.values[3] <- t.Topping.H03$p.value


# adjusting p-values with "bonferroni" method
adjusted.p.values <- p.adjust(unadjusted.p.values, method = "bonferroni")
adjusted.p.values

## Which is nothing but ....
unadjusted.p.values * 3

# T-test for Butter ===================================
# For Butter : No adjustment needed -----------------------------
t.Butter.H01 <- t.test(dataset$Ants[dataset$Butter == "with"], 
       dataset$Ants[dataset$Butter == "without"], 
       alternative = "two.sided", paired = FALSE, var.equal = TRUE)

t.Butter.H01$p.value


```

